Cultural Transmission and the Social Contexts of Technological Transitions during the Late Paleolithic of Korea

### 2023-03-09 analysis/paper/paper.qmd
Add Crema's new paper

To measure transmission biases in tool-making, we draw on Darwinian evolutionary theory. This body of theory has aided archaeologists in understanding a variety of technological innovations and related human behaviors [@dunnell1980evolutionary; @richerson1992cultural; @lipo1997population; @bettinger1997evolutionary; @bettinger1999point; @eerkens2008cultural; @mesoudi2008cultural]. Cultural evolutionary approaches have been effective at explaining cultural changes using mechanisms of inheritance, variation-generating processes, and selection. These mechanisms are similar to biological evolution, but in cultural systems they have unique and distinctive properties [@Creanza_Kolodny_Feldman_2017; @Claidiere_Scott_Phillips_Sperber_2014; @Mesoudi_Whiten_Laland_2004; @Acerbi_Mesoudi_2015]. Social transmission strongly affects these key evolutionary processes [@Whiten_2017] and can occur through various learning processes, such as stimulus enhancement, emulation, imitation, and teaching [@o2000applying; @schillinger2014copying]. We focus here on how different social contexts can result in different modes of social transmission of tool-making skills among hunter-gatherer populations.

Cultural transmission theory holds that information about social and technological behaviours is acquired through interaction with other individuals and the environment [@boyd1988culture; @richerson1992cultural]. Individuals learn by themselves (e.g. trial and error) or from each other by sharing information. Information can be modified (also known as 'biased') depending on an individual's transmission context and cultural repertoire [@richerson2008not]. Modification of information can occur by recombination, loss, or partial alteration [@eerkens2005cultural; @o2017dual]. These transmission biases can be important loci of changes in material culture, and can be influenced by the social contexts of cultural transmission [@heyes1994social; @eerkens2007cultural; @lycett2015cultural; @creanza2017cultural; @kendal2018social]. These biases include guided variation (where individuals learn new behaviors through trial and error), content-based bias (where some aspect of the transmitted instructions, such as cultural preferences, makes them more likely to be adopted), frequency-based bias (where an individual is biased to choose particular behaviors based on their perceived frequency in the population, such as extremely popular or rare behaviors), and indirect bias (where a behavior is transmitted because of its association with other attributes, such as the prestige or skill of other individuals) [@boyd1988culture; @richerson1992cultural; @o2017dual; @richerson2008not].

>>>>

To measure transmission biases in tool-making, we draw on Darwinian evolutionary theory. This body of theory has aided archaeologists in understanding a variety of technological innovations and related human behaviors [@dunnell1980evolutionary; @richerson1992cultural; @lipo1997population; @bettinger1997evolutionary; @bettinger1999point; @eerkens2008cultural; @mesoudi2008cultural; @crema2023cultural]. Cultural evolutionary approaches have been effective at explaining cultural changes using mechanisms of inheritance, variation-generating processes, and selection. These mechanisms are similar to biological evolution, but in cultural systems they have unique and distinctive properties [@Creanza_Kolodny_Feldman_2017; @Claidiere_Scott_Phillips_Sperber_2014; @Mesoudi_Whiten_Laland_2004; @Acerbi_Mesoudi_2015]. Social transmission strongly affects these key evolutionary processes [@Whiten_2017] and can occur through various learning processes, such as stimulus enhancement, emulation, imitation, and teaching [@o2000applying; @schillinger2014copying]. We focus here on how different social contexts can result in different modes of social transmission of tool-making skills among hunter-gatherer populations.

Cultural transmission theory holds that information about social and technological behaviors is acquired through interaction with other individuals and the environment [@boyd1988culture; @richerson1992cultural]. Individuals learn by themselves (e.g. trial and error) or from each other by sharing information. Information can be modified (also known as 'biased') depending on an individual's transmission context and cultural repertoire [@richerson2008not]. Modification of information can occur by recombination, loss, or partial alteration [@eerkens2005cultural; @o2017dual]. These transmission biases can be important loci of changes in material culture, and can be influenced by the social contexts of cultural transmission [@heyes1994social; @eerkens2007cultural; @lycett2015cultural; @creanza2017cultural; @kendal2018social]. These biases include guided variation (where individuals learn new behaviors through trial and error), content-based bias (where some aspect of the transmitted instructions, such as cultural preferences, makes them more likely to be adopted), frequency-based bias (where an individual is biased to choose particular behaviors based on their perceived frequency in the population, such as extremely popular or rare behaviors), and indirect bias (where a behavior is transmitted because of its association with other attributes, such as the prestige or skill of other individuals) [@boyd1988culture; @richerson1992cultural; @o2017dual; @richerson2008not].
